\chapter{Transcriptome Variation Under Novel Growth Conditions}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%                         Methods                                   %%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Methods}

\subsection{Growth cabinet control software}

The Research School of Biology has acquired computer-controlled growth cabinets with multispectral
LED light arrays. These cabinets are digitally controlled growth cabinets (Conviron ; %<+model+>
Conviron, Winnepeg, Canada) equipt with multispectral LED growth lamps (Heliospectra Model L4A Series 10;
Heliospectra AB, Sweden) and image-based phenomics systems (in-house solution). This suite of
equiptment is all computer controllable, however no interface to control these systems as a whole
exists. Thus, I have created software which is able to control these cabinets, enabling the creation
of novel growth conditions.

The software I have created, \texttt{spcControl}, gives control commands to both growth cabinets and
retrofitted LED light lamps. This software is implemented in the Python programming language, and
allows conditions generated in a defined format by external software to be parsed, converted to
control commands, and sent in concert to all components of the SpectralPhenoClimatron. It runs on a
computer situated next to the chambers, and can control many chambers at once <++>


\subsection{Design of novel growth conditions}

Novel growth condiitons were designed to emulate some charachteristics of growth conditions
experienced by plants in natural environments. Using SolarCalc <+CN+>, I have created simulated
climates derived from historical observations and models of the climate of the region surrounding
Temora, New South Wales. This is a trial location for various agricultural crops, and is in the same
climate zone as much of the native range of \textit{Arabidopsis thaliana}. SolarCalc has an large
number of parameters which can be set. The settings I used in the creation of the novel conditions
are described in \autoref{tab:biol-meth-solarcalcsettings}. Solarcalc was run with these model
parameters, to produce a model of growth conditions which simulate elements of a natural
environment.

\begin{table}
  \centering
  \begin{singlespace}
    \begin{tabular}{lc}
      Parameter & Temora\\
      \hline
      Simulation End Date & 31/12/2012 \\
      Simulation Start Date & 01/09/2012 \\
      Chamber Max RH & 85 \\
      Chamber Min Temp & 5C \\
      Chamber Date & 01/03/2013 \\
      Date Format & US \\
      Site Elevation & 300m \\
      Site Latitude & -34.446556 \\
      Site Longitude & 147.53334 \\
      Timezone & Sydney \\
      Update Frequency & 1 min \\
      Weather Year & 2010 \\
      \hline
      Lighting & LED \\
      Set to Threshold & Yes \\
      LED1 Power Threshold & 5.00\% \\
      LED1 Wavelength & 400nm \\
      LED1 Weight Multiplier & 5.2 \\
      LED2 Power Threshold & 5.00\% \\
      LED2 Wavelength & 420nm \\
      LED2 WeightMultiplier & 4.12 \\
      LED3 Power Threshold & 5.00\% \\
      LED3 Wavelength & 450nm \\
      LED3 Weight Multiplier & 4 \\
      LED4 Power Threshold & 5.00\% \\
      LED4 Wavelength & 530nm \\
      LED4 Weight Multiplier & 3.92 \\
      LED5 Power Threshold & 5.00\% \\
      LED5 Wavelength & 630nm \\
      LED5 Weight Multiplier & 4.88 \\
      LED6 Power Threshold & 5.00\% \\
      LED6 Wavelength & 660nm \\
      LED6 Weight Multiplier & 0.68 \\
      LED7 Power Threshold & 5.00\% \\
      LED7 Wavelength & 740nm \\
      LED7 Weight Multiplier & 3.64 \\
      \hline
      AOD & 0.27 \\
      Azimuth & 180 \\
      Slope & 10 \\
      Soil Albedo & 0.5 \\
      Solar Tilt & 3.5 \\
      Solar Constant & 1360 \\
      Track Sun & False \\
      \hline
    \end{tabular}
  \end{singlespace}
  \caption{Parameter settings of SolarCalc used in the creation of novel conditions.}
  \label{tab:biol-meth-solarcalcsettings}
\end{table}

SolarCalc simulates the climate of a location without any weather, however I aimed to create
condition which mimic cloudy and intermittently cloudy days. Thus, to create such conditions,
post-processing work is requred. Firstly, SolarCalc was re-run with a neutral density shade
parameter of 70\%, meaning that the intentisy of each LED channel was 70\% that of the original
model, forming the shaded model, which aimed to emulate a cloudy day. Then, the result of both the
original model, and this shaded model were spliced together to form a condtion whose light intensity
changed between shaded and original models on a two hour shaded, one-hour original rotation,
creating the fluctating condition designed to emulate an intermittently cloudy day. The temperature,
humidiy and light quality of these conditions remained unchanged. These additional condtiions allow
us to compare the effect of different light regimes within the framework of such pseudo-natural
dynamic growth conditions which SolarCalc creates. 
\todo{This sounds too much like what got destroyed in the initial examiners meeting}







%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%                         Results                                   %%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Results}



