%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%                       Discussion                                  %%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Discussion}
\label{chap:dis}

The study of plant-environment interactions is crucial to improving agricultural yield and
understanding functional plant ecology in a changing climate. The study of mechanisms by which
plants tolerate or respond to  abiotic stresses is of particular importance, as abiotic stresses
cause many hundreds of billions of dollars of crop losses and incalculable environmental damage
\autocite{mittler_abiotic_2006}. Several recent reviews of abiotic stress research have suggested a
changed focus for research, away from studying stresses individually towards more holistic
combined stress studies \autocite{mittler_abiotic_2006, mittler_genetic_2010}. Plants in ecological
or agricultural settings commonly experience environments unfavourable to optimal growth, composed
of multiple abiotic (and biotic) stresses \autocite{boyer_plant_1982, mittler_abiotic_2006}. 
Moreover, natural environments are dynamic, and challenge plants to rapidly adapt to heterogeneous
environmental conditions on a daily basis. Laboratory
study of these stresses should therefore account for recent findings suggesting that interactions
between abiotic stresses in field conditions are not simply additive, rather are influenced by the
dynamic combinations of stresses their environments impose upon them \autocite{mittler_abiotic_2006,
atkinson_interaction_2012}.

This thesis has developed methods to reproducibly induce abiotic stresses in laboratories in 
combinations analogous to those in which they can appear naturally and assay plant growth under such
conditions, analyse plant response to responses, and presents a preliminary proof-of-concept transcriptomics
dataset analysing light response in the context of combinatorial abiotic stresses.


\section{Novel Dynamic Growth Conditions}
\label{sec:dis-dc}

Plant biologists often conduct laboratory experiments in controlled growth facilities. These
facilities may lack the ability to mimic characteristics of their natural environment, as they are
rarely built from hardware enabling an investigator to do so. While static growth conditions used in
laboratories are sufficient to uncover core mechanisms of plant responses to stress, this is not
always the case. In fact, evidence from field studies has uncovered phenotypes that are cryptic
under static laboratory conditions, and that may be observed and studied closely under dynamic
laboratory conditions \autocite{kulheim_rapid_2002, mittler_abiotic_2006, mishra_arabidopsis_2012,
wituszynska_multivariable_2013}. By creating software which connects existing but disparate
technologies, I have enabled the implementation of growth conditions which can mimic regional
climates or weather. The combination of computer controllable growth cabinets fitted with
multi-spectral LED arrays allow for experimentation to consider important parameters that vary in
natural regional climates, including temperature, humidity, and light quality and quantity.

In creating dynamic growth conditions, I aimed to combine some elements of the climates experienced
by plants in nature with the reliability, reproducibility and convenience of laboratory growth
conditions. Diurnal trends in light intensity, temperature and humidity follow those observed in the
recent past at the model location \autocite{bureau_of_meteorology_climate_2013}. However, daily
minima in temperature and maxima in humidity are tempered by hardware limitations. Similarly, daily
integrated light intensity is lower due to limitations in the brightness of LED arrays used in the
SpectralPhenoClimatron. The light produced by LED arrays does not cover the photosynthetically
active portion of the visible and adjacent spectrum with an even intensity of light per wavelength
as sunlight does, although LED arrays do so with broader peaks of increased spectral intensity than
fluorescent or incandescent lamps. These are important limitations to the SpectralPhenoClimatron,
and addressing them in future work is a priority. Circannual variation in temperature and humidity
follow similar climactic trends as the historical mean observations of weather at the trial
location, with temperature slowly increasing, humidity decreasing and photoperiod lengthening,
however light intensity does not gradually increase in the same manner as the observed climate.

While the dynamic growth conditions implemented in the SpectralPhenoClimatron mimic elements of
natural environments, they are not designed to be accurate representations of conditions encountered
in nature. In nature, weather provides a layer of stochasticity upon the broad trends in climate
\autocite{citation needed}. However, the developments I have described in this thesis facilitate the
emulation of stressful growth environments which approximate weather-induced abiotic stresses plants
experience in the natural environments. The system I have developed is not only applicable to
studies discussed in this thesis, and has been used in experiments beyond the scope of this thesis
including virtual reciprocal transplants and gene-by-environment interaction QTL mapping. The study
of standing genetic variation within regional climates allows links between genotype and reaction to
some environmental parameter, and can shed light on mechanistic links via methods including GWAS
\autocite{li_genetics_2006, li_association_2010, brachi_genome-wide_2011}. However, when modelling
plant growth in, or reaction to, environmental conditions, the system I have developed allows for
non-stochastic weather to be imposed over the climatic trends (e.g. the fluctuations in light
modelling intermittent cloud cover used in this thesis).

\todo{mention ecology, ``virtual
reciprocal transplants'' etc} \todo{reiterate reliability, and make it first person} \todo{CRUCIAL-
need a link}


\section{Improved analytic methods for RNAseq}
\label{sec:dis-bioinf}

RNAseq is a precise method to quantify genome-wide expression
\autocite{wang_rna-seq:_2009}. RNAseq can reveal hidden phenotypes and subtle environmental effects
of a plant's growth environment \autocite{martin_catalyzing_2013}, giving insights into development,
regulatory mechanisms, signalling pathways, acclimation and many other aspects of plant biology. It
is sufficiently sensitive to measure the subtle differences in expression between both closely
genetically related organisms (à la expression QTL mapping, \textcite{sun_eqtl_2013}), and subtle
environmental effects (e.g. differential response to dynamic growth conditions).

Reproducibility, accuracy, and performance of the computational analysis of any dataset is crucial.
Inaccurate and poorly reproducible analyses have lead to embarrassing errors and retractions in many
fields \autocite{peng_reproducible_2009, herndon_does_2013}.  Computational performance of analysis
software is central to interactive analysis of datasets; enabling fast analysis of datasets allows
researchers to explore their data without the requirement for expensive clusters or supercomputers.
The design and implementation of the RNAseq analysis framework I have created specifically address
reproducibility, accuracy, and performance, by selecting the latest advances in high-performance,
accurate analysis software, and providing a simple structure to an analysis that allows exact
reproduction of the entire analysis.

%summarise and discuss improvements to all 3. cite reproducible research/software carpentry type
%papers here. This should be split into 2 paragraphs I think)

The specific algorithms used in analysis of RNAseq datasets is a field of active technical research.
Validation of pipelines designed around the Subread aligner \autocite{liao_subread_2013} reiterate
the software authors' claims of increased speed and accuracy. A recent publication suggests superior
performance of the trimmed mean of M values normalisation method proposed by
\textcite{robinson_scaling_2010} and implemented by \textcite{robinson_edger:_2010} compared to its
competitors \autocite{rapaport_comprehensive_2013}. However, this review did not consider recently
published statistical techniques such as QLspline \autocite{lund_detecting_2012}. Further
investigation of the state of the art in RNAseq statistical and computational analysis software is
of great importance to studies of differential expression, and the framework presented in this
thesis is specifically designed for modularity, enabling substitution of components for improved
versions or alternatives with minimal effort.

\todo{mention txome subtlety}

%(describe its modularity and possible ``future directions'', i.e. adding support for bigger and
%better programs. perhaps back in technical discussion)

When transcriptome variation within a study system is subtle, sensitive and accurate methods are
required to glean information. This includes accurate analysis software (see above) and optimal
experimental design. I examined the effect of sequencing depth on power to detect differential
expression, and found that below 5 million reads per sample, power to detect differential expression
with three replicates diminishes rapidly (see \autoref{fig:br-res-detagsvscoverage}). Several recent
studies and reviews suggest that the current informal standard of RNAseq experimental design, which
emphasises sample sequencing coverage over replication, is unwise.
\textcite{rapaport_comprehensive_2013} find increasing the level of replication over sequencing
depth yields more differentially expressed genes, while \textcite{kliebenstein_exploring_2012}
demonstrate that, even when sequence coverage was very low, nearly all expression QTLs could be
mapped. In light of these data, my analysis of sequencing depth is conservative, as it does not
consider the increases in replication made possible by increasing the number of samples sequenced
per unit cost by a factor of two or more.

\todo{Biological variation control and multi environment treatments are needed to provide specificity
of changes.  Gene Atlas idea.}

\section{Elucidating Response to Light Intensity Under Dynamic Growth Conditions}
\label{sec:dis-txvar}

The transcriptomes of \textit{Arabidopsis} grown under dynamic growth conditions have been
observed. A preliminary RNAseq dataset characterised by high levels of biological noise indicated
differential expression of hundreds of transcripts between growth conditions and in response to
treatment with hot excess light. QPCR analysis showed increased steady-state expression of know
stress responsive genes under dynamic growth conditions, and reduced fold-change induction of
stress genes upon application of hot excess light. Together, these data indicate patterns of
differential expression observed under dynamic growth conditions, and provide limited and tentative
support for the hypothesised hardening of plants to excess light.
\todo{this paragraph needs a lot of thought, maybe move this one to summary of ch4, and rewrite the
one that ends up here}

The large amount of biological noise and resulting lack of power to detect differential expression
is likely caused by a combination of factors. A rapid (within minutes) reduction in transcript
abundance of hot excess light induced genes upon removal from hot excess light treatment has been
observed (pers.  comm. Peter Crisp). In light of these data, the high temporal error in sampling
which exists in the dataset that I created may explain some of the biological variability.
Additionally, as plants were of an advanced age (5 weeks) when samples were taken rapid estimation
of the particular leaf that was samples was infeasible, thus the largest expanded leaf was taken.
This variation in leaf number may be an additional reason for the high variability in this dataset,
as the transcriptome is dependent on leaf developmental stage \autocite{gordon_systemic_2013,
carmody_rapid_2013}. A final possible explanation for the high biological noise is variation in
harvesting time. While plants were harvested as quickly as possible, each replicate took 3 hours to
harvest. Previous studies have found circadian effects to both the general transcriptome
\autocite{covington_global_2008, ptitsyn_comprehensive_2008} and in transcriptional response to
biotic stress \autocite{wilkins_time_2010}. The harvesting techniques and experimental design utilised in
this study were suited to experiments that mapped eQTLs for stress-responsive genes. For simple
detection of differential expression however, they were not optimal and likely contributed to the
high level of biological noise in the obtained dataset. Ideally, true internal controls grown under
static growth conditions, and higher levels of biological replication would be more appropriate.
These shortcomings result in reduced statistical power, and are a possible cause for the lower than
expected number of genes whose expression was induced or repressed upon treatment with hot excess
light.

Given this high level of biological noise, all further analyses of differential expression should be
approached with caution, even if statistical techniques are sufficiently advanced to allow detection
of differential expression. Despite this, patterns of expression similar to those observed by other
authors have been noted in this study. Hundreds of genes were differentially expressed in response
to one hour of excess light in each dynamic light condition, a similar magnitude to previous
examinations of similar stresses \autocite{rossel_global_2002, rossel_systemic_2007,
gordon_systemic_2013, kimura_identification_2003}. Specifically, the
induction of heat shock family proteins observed by \textcite{rossel_global_2002} was also observed
here. Additionally, gene ontology (GO) term analysis indicates commonality of the transcriptional
response to hot excess light with heat and oxidative stresses, as observed in previous studies
\autocite{rossel_global_2002}. Together, these data suggest that the response of plants grown under
dynamic light conditions to hot excess light is similar to that of plants growth under static light
conditions, albeit with somewhat tempered induction of hot excess light-responsive marker genes. An
understanding of the power needed and variation present in a typical experiment has been identified.

Limited transcriptomic evidence of the effect of light intensity within the framework of dynamic
light conditions was observed. Gene ontology (GO) term enrichment provides evidence that plant
response to hot high light is preserved in plants grown under dynamic growth conditions hypothesised
to induce acclimation to high light or fluctuations in light intensity
(\autoref{fig:txv-res-govenn}). Additionally, GO term enrichment analysis details the overlap and
interaction with other abiotic stresses. GO terms  enriched in genes with differential steady-state
expression in plants grown under sufficient light dynamic growth conditions and excess or fluctuating
light dynamic growth conditions exhibit limited overlap with those observed in previous studies of
acute changes in light intensity, as hypothesised. However, this evidence of hardening is contrasted
by the number of genes differentially expressed in response to hot high light under dynamic growth
conditions. If acclimation to modest excesses in light intensity result in reduced fold-change
induction of stress-inducible genes, as qPCR quantification of APX2 expression suggest, then either
the relative order of hardening hypothesised is incorrect, or the level of biological noise in this
dataset prevented observation of such phenomena.
\todo{fix up go paragrahph}

\todo{add model of mop vs bucket}

A more detailed analysis of this dataset may be warranted by these findings. In particular,
examination of possible harvesting or treatment block effects may, if any effect exists, reduce the biological coefficient
of variation \autocite{robinson_edger:_2013}. A detailed analysis of all genes called as
differentially expressed may be of use, as high biological noise may not affect the estimates of
differential expression in all genes. I hypothesise that upon doing so, genes previously found to be
differentially expressed between lab and field growth conditions would have differential
steady-state expression between sufficient, fluctuating and excess light conditions, with expression
being highest in the excess and lowest in sufficient light conditions. 
%\begin{itemize}
%  \item check for block effect
%  \item look at detail of individual genes
%  \item Hypothesise as to what good dataseet would find
%\end{itemize}
\todo{ensure expt. methods stuff here makes it into ch4 methods}

Study of the transcriptome is an established method to assess the responses of plants to their
environment, but it is far from the only method of doing so. Analysis of chlorophyll fluorescence of
plants acclimating to fluctuating light demonstrated up-regulation of NPQ
\autocite{alter_acclimatory_2012, gordon_systemic_2013}, and study of such characteristics under
dynamic light conditions is ongoing, but beyond the scope of this thesis. Analysis of accumulation
of stress metabolites has elucidated metabolites involved in acclimation to combined stresses in
field conditions \autocite{jankanpaa_metabolic_2012}, warranting similar analysis of metabolites in plants
grown under dynamic growth conditions. However, transcriptomics remains a sensitive, accurate and
useful tool to measure plants' responses to their environment, particlarly in a systems biology
approach that integrates metabolomic, phenomic, transnscriptomic and physiological responses to
applied stresses.


\section{Discussion of data in broader context of abiotic stresses}
\label{sec:dis-combstress}
\todo{better title for this section}

\todo{intro to section interpretation discussion}

While the effect of altered light intensity on plants and their transcriptomes has been studied in
depth, few authors have studied under laboratory conditions analogous to those to which plants are
adapted.  Mutants in pathways important in survival or fecundity of plants in ecological or
agricultural settings may show little or no detrimental phenotype under benign laboratory conditions
\autocite{kulheim_rapid_2002}. This further underlies the importance of examining the physiological,
metabolomic and transcriptional responses of plants to field-like combinations of stresses
\autocite{jankanpaa_metabolic_2012, mishra_arabidopsis_2012, wituszynska_multivariable_2013}. By
examining the responses of plants to their environment in conditions similar to those which they
have evolved, and studying stress response to field-like combinations of stresses, it may be
possible to obtain a more direct picture of the role of parts of the genome with little function in
the artificially benign conditions under which many laboratory experiments on plants are conducted.
\todo{this could do with more refs}

Studying the effect of hardening on the transcriptional response of plants to stress may reveal
mechanistic insights of response genes. As plants acclimate to altered growth conditions,
alteration of steady-state transcription often occurs \autocite{hihara_dna_2001, alter_acclimatory_2012,
heinrich_transcriptomic_2012}. Transcriptome profiling during acclimation to cold stress,
\autocite{fowler_arabidopsis_2002, chawade_putative_2007}, excess light
\autocite{gordon_systemic_2013,page_influence_2012}, and drought \autocite{ding_multiple_2012} reveals a
transcriptional response to long-term stress exposure. The elucidation of genes underlying such
acclimatory responses would yield insights into mechanisms by which plant tolerance of chronic
abiotic stress could be improved.
\todo{need more refs and less fluff for this paragraph}
%(my hypothesis RE ``mop and bucket'' vs ``guardian'' genes.)


Scientists have repeatedly developed crop lines tolerant to stress assays ``in the lab'' that, when
trialled under agricultural condition, either do not show stress tolerance or have increased
susceptibility to other stresses or combinations of stresses \autocite{mittler_abiotic_2006,
mittler_genetic_2010, atkinson_interaction_2012, wituszynska_multivariable_2013}. Some of the
difficulty in translation of stress tolerance from lab to field can be explained by detrimental
interaction of stresses \autocite{mittler_genetic_2010}. By performing genetic screens or genome
wide association studies for stress tolerance under dynamic growth conditions, which mimic
field-like combinations of stresses, it may be possible to elucidate mechanisms of tolerance to
specific stresses which do not conflict with mechanisms of tolerance to other abiotic stresses, as
tolerance is assayed in a mildly stressful environment that mimics that encountered under
agricultural conditions.

\todo{first sentence needs rewording}

\section{Future Directions}
\label{sec:dis-futdir}

The emergence of transcriptional patterns despite a dataset with high noise warrants further
investigation. Brief examination of fold-change values indicates that biological noise is the
primary factor for reduced differential expression. Utilising recent research on the underlying
biology of transcriptional response (Peter Crisp, unpublished data) and 
RNAseq molecular methodology \autocite{kumar_high-throughput_2012} and experimental design
\autocite{rapaport_comprehensive_2013}, a repeated experiment which incorporates faster sampling and
more accurate application of hot excess light stress, along with many more biological replicates
(5-8 replicates) per condition, may further elucidate the patterns of differential expression under
dynamic growth conditions, and determine the effect of light intensity on the transcriptome
under dynamic light conditions. 
\todo{rework this a bit}

Expression QTL mapping presents an opportunity to simultaneously discover novel transcriptional
patterns, and their underlying regulatory mechanisms \autocite{keurentjes_regulatory_2007}. Samples
to perform a limited eQTL mapping experiment were collected but not processed due to constraints on
time and statistical power. Once genetic material suitable for such analyses has been obtained,
successfully conducting an analysis to find gene-by-environment interactions in and eQTLs
controlling differential expression of transcripts in response to altered light intensity within the
framework of dynamic growth conditions would elucidate not only response mechanisms, but their
underlying regulators. This information could be used by crop improvement programs to select for
lines with improved ability to respond to or tolerate field-like combinations of stresses.

Application of real-time phenomic analysis to plants grown under dynamic growth conditions, combined
with QTL mapping or GWAS analyses, may provide novel insights into physiological responses to
combinations of abiotic stresses. Phenomic traits that correlate with 
transcriptional or physiological response to or tolerance of combinations of abiotic stresses
experienced under dynamic growth conditions should be selected. Such traits may include chlorophyll fluorescence
measurements of redox state or NPQ \autocite{alter_acclimatory_2012, kulheim_rapid_2002}, or visual
estimation of pigmentation caused by anthocyanin accumulation in response to photooxidative stress.
GWAS and QTL mapping for these non-destructive measures could be conducted concomitantly with
mapping of eQTLs discussed above. Such experiments would present integrative results which give
holistic insight to mechanisms of tolerance of or response to combinations of abiotic stresses.


\section{Conclusions}
\label{sec:dis-conclusion}

This thesis has developed several methods to study abiotic stresses in the framework of dynamic
conditions, and presents a preliminary dataset which examines plant response to such conditions.
Aiming to design and implement dynamic growth conditions that mimic regional climates, within this
thesis I have presented both software and protocols to implement such conditions, and have created
conditions to test the effect of light intensity in combination with abiotic stresses. Aiming to
select optimal software for the high-throughput study of transcriptome dynamics using
High-throughput Sequencing, and implement a framework for generation of analysis pipelines to do so,
I have presented a comprehensive framework for the creation of pipelines applicable to experiments
beyond those discussed in this thesis. Additionally, I have, using this framework, implemented
pipelines to analyse generic RNAseq datasets. Finally, aiming to determine the transcriptional
response of \textit{Arabidopsis thaliana} to the combinatorial application of abiotic stresses using
dynamic growth conditions, I created a preliminary RNAseq dataset, and samples for preliminary eQTL
and phenomic QTL mapping datasets. This RNAseq dataset exhibited high biological noise, preventing
reliable detection of differential expression on similar scales to similar published experiments,
however yielded insight into patterns of differential expression that warrant further investigation.
This work lays the foundation for further work elucidating the response to and tolerance of field-like
combinations of abiotic stresses in a manner compatible with that suggested by recent reviews
\autocite{mittler_abiotic_2006, atkinson_interaction_2012}, and for the further discovery of the genetic
architecture underlying any tolerance or response discovered.
\todo{add juice to final sentence}
