\chapter{Design and Implementation of Dynamic Growth Conditions}
\chaptermark{Dynamic Growth Conditions}
\label{chap:dc}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%                     Background & Aims                             %%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\section{Aims and Hypothesies}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%                         Methods                                   %%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Methods}

\subsection{Growth cabinet control software}

The Research School of Biology has acquired computer-controlled growth cabinets with multispectral
LED light arrays. These cabinets are digitally controlled growth cabinets (Conviron PGC20 reach-in
Chambers; Conviron, Winnepeg, Canada) equipt with multispectral LED growth lamps (Heliospectra Model
L4A Series 10; Heliospectra AB, Sweden) and image-based phenomics systems (in-house solution). This
suite of equiptment is computer controllable, however no interface to control these systems as a
whole exists. Thus, I have created software which is able to control these cabinets, enabling the
creation of dynamic growth conditions.

The software I have created, \texttt{spcControl}, gives control commands to both growth cabinets and
retrofitted LED light lamps. This software is implemented in the Python programming language, and
allows conditions generated in a defined format by external software to be parsed, converted to
control commands, and sent in concert to all components of the SpectralPhenoClimatron. It runs on a
computer situated next to the chambers, and can control many chambers at once <++>


\subsection{Design of dynamic growth conditions}
\label{subsec:txv-mth-growthcond}

Novel growth condiitons were designed to emulate some charachteristics of growth conditions
experienced by plants in natural environments. Using SolarCalc \autocite{spokas_estimating_2006}, I
have created simulated climates derived from historical observations and models of the climate of
the region surrounding Temora, New South Wales. This is a trial location for various agricultural
crops, and is in the same climate zone as much of the native range of \textit{Arabidopsis thaliana}.
SolarCalc has an large number of parameters which can be set. The settings I used in the creation of
the dynamic conditions are described in \autoref{tab:dc-mth-solarcalcsettings}. Solarcalc was run
with these model parameters, to produce a model of growth conditions which simulate elements of a
natural environment.

\kmdefaulttable
{dc-mth-solarcalcsettings}
{./tables/bio-mth-solarcalcsettings.csv}
{lc}
{Parameter & Temora Setting}
{\param=Parameter, \tem=Temora}
{\param & \tem}
{Parameter settings of SolarCalc used in the creation of dynamic conditions.}

SolarCalc simulates the climate of a location without any weather, however I aimed to create
condition which mimic cloudy and intermittently cloudy days. Thus, to create such conditions,
post-processing work is requred. Firstly, SolarCalc was re-run with a neutral density shade
parameter of 70\%, meaning that the intentisy of each LED channel was 70\% that of the original
model, forming the shaded model, which aimed to emulate a cloudy day. Then, the result of both the
original model, and this shaded model were spliced together to form a condtion whose light intensity
changed between shaded and original models on a two hour shaded, one-hour original rotation,
creating the fluctating condition designed to emulate an intermittently cloudy day. The temperature,
humidiy and light quality of these conditions remained unchanged. These additional condtiions allow
us to compare the effect of different light regimes within the framework of such pseudo-natural
dynamic growth conditions which SolarCalc creates.
\todo{This sounds too much like what got destroyed in the initial examiners meeting}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%                         Results                                   %%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results}
\label{sec:dc-results}

\subsection{Sufficient, Excess and Fluctuating Dynamic Growth Conditions}



\subsection{Novel Dynamic Growth Conditions That Simulate a Continental Gradient}


