%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%                       Introduction                                %%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Introduction}

Nearly all terrestrial biomass, including human life, depends on the primary productivity of higher
plants. This ability is reduced when plants are stressed by interaction with their environment.


%the quality or quantity of light a plant receives is not optimal. Excess light is particularly
%damaging, causing both reductions in photosynthetic ability (photoinhibition) and cellular or
%tissue damage and death. Thus, plants have evolved mechanisms by which the detrimental effects of
%excess light can be minimised. These mechanisms, collectively termed photoprotection, work to
%dissipate excess energy, reduce the amount of light absorbed, or prevent or repair any damage
%caused \autocite{niyogi_photoprotection_1999}.
%\todo{Get some measure of the global cost of photodamgage}
%\todo{make this less mechanistic, more borevitzy - subtle changes is the key}

%  Although difficult to quantify on a global scale, the
%  cost of the  detrimental effects of excess light on primary productivity is high
%  \autocite{raven_cost_2011}.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   Stress and Coping Mechanisms   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Abiotic Stress: A limit to Plant Productivity}
\label{sec:intro-stress}

Abiotic stresses are those applied to plants from their growth environment. This includes
environmental variables including temperature, humidity, salinity and water availability, and light
quality and quantity amongst many others. (describe specific stresses and their general effects)

Plant productivity is limited by abiotic stresses, as their detrimental effects affect the ability
of plants to assimilate biomass, and coping mechanisms plants have evolved generally come at a yield
penalty. (cite paper/review describing effects, paper showing constitutive expression of some HSF, or
other stress mitigator)


\subsection{Photooxidative damage and Photoprotection: Coping with Detrimental Effects of Excess Light}
\label{subsec:intro-stress-photodamage}

Plant primary productivity depends on their ability to utilise energy from light to fix inorganic
carbon into organic biomass. This ability is reduced when the quality or quantity of light a plant
receives is not optimal. Excess light is particularly damaging, causing both reductions in
photosynthetic ability (photoinhibition) and cellular or tissue damage and death.
(\autocite{niyogi_photoprotection_1999}, other reviews of light stress and Photooxidative damage)

The Photosynthetic machinery, upon exposure of plant to excessive light, produced harmful
byproducts. These include reactive oxygen species which damage both the photosynthetic machinery and
cellular components in general. (summary of ROS production after EL, photooxidative stress,
\autocite{asada_production_2006, suzuki_ros_2012} and other papers)

Because of the indispensability of photosynthesis, plants have evolved 
evolved mechanisms by which the detrimental effects of excess light can be minimised. These
mechanisms, collectively termed photoprotection, work to dissipate excess energy, reduce the amount
of light absorbed, or prevent or repair any damage caused \autocite{niyogi_photoprotection_1999}.
(summary and analysis of \autocite{niyogi_photoprotection_1999, takahashi_photoprotection_2011})

Several mechanisms of photoprotection occur within the chloroplast. Non-photochemical quenching
(NPQ) dissipates excess energy from excited state chlorophyll molecules as heat
\autocite{muller_non-photochemical_2001}. It occurs in photosystem II , and is particularly
important during rapid changes in intensity \autocite{kulheim_rapid_2002}. Cyclic electron flow
occurs in photosystem I and acts by decreasing the pH of the thylakoid lumen, which in turn is
thought to stabilise the oxygen evolving complex and aid NPQ \autocite{takahashi_how_2009}. State
transitions act by phosphorylation of photosystem II and light harvesting complex II proteins.
State transitions reversibly alter the balance of excitation energy between photosystems I and II
\autocite{tikkanen_state_2006}. Through reversible dissociation of the light harvesting complex II
and photosystem II , state transitions lower the light harvesting ability of photosystem II  and
prevent absorption of excess light by photosystem II  \autocite{johnson_photoprotective_2011}. The
mutants \textit{\underline{N}ON-\underline{P}HOTOCHEMICAL \underline{Q}UENCHING \underline{1}}
(\textit{NPQ1}) and \textit{\underline{N}ON-\underline{P}HOTOCHEMICAL \underline{Q}UENCHING
\underline{4}} (\textit{NPQ4}), \textit{\underline{P}ROTON \underline{G}RADIENT
\underline{R}EGULATOR \underline{5}} (\textit{PGR5}) and \textit{\underline{S}TATE
\underline{T}RANSITIO\underline{N} \underline{7}} (\textit{STN7}) and \textit{\underline{S}TATE
\underline{T}RANSITIO\underline{N} \underline{8}} (\textit{STN8}) are defective in these three
mechanisms respectively.


\subsection{Abiotic Stresses: Studied Individually, Encountered Combinatorially}
\label{subsec:intro-stress-combstress}

Despite the large body of published works elucidating coping mechanisms of and responses to several
abiotic stresses, the majority of works focus on an abiotic stress in isolation
\autocite{atkinson_interaction_2012, mittler_abiotic_2006}. However, in natural or agricultural
situations, plants rarely experience an abiotic stress in isolation. This has lead to difficulty in
translating stress-tolerant lines of plants developed in labs to more hardy field lines
\autocite{mittler_abiotic_2006}.

Field-grown plants tend to demonstrate reduced survival, reproduction and altered transcriptional
responses compared to those grown in comparable lab conditions. For example, in Arabidopsis thaliana
the reduction in reproductive success caused by the \textit{npq1} and \textit{npq4} mutants was
severe in the field but had no impact under constant light in the lab \autocite{kulheim_rapid_2002}.
Photoinhibition, physiological symptoms of abiotic stress and expression of high light-induced
transcripts were more severe in field-grown \textit{Solidago altissima} than those grown in
comparable lab conditions \autocite{barua_interactive_2006}.

Temperature interacts with light stress in a detrimental fashion \autocite{mittler_abiotic_2006}.
At cold temperatures, unacclimated plants are less able to dissipate excess energy to avoid
Photooxidative damage (\autocite{armstrong_assessing_2007}, probably need a better paper, check their refs.
Also talk about hot vs cool EL, from \autocite{gordon_systemic_2013}).

\todo{Alot more on field grown plants - 2 pgf}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   Responses   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Plant Responses to Abiotic Stresses}
\label{sec:intro-resp}

Plant responses to abiotic stress have been closely studied recently using a variety of techniques.
(outline some of the techniques used: chlorophyll fluorescence, transcriptomics (qPCR, microarray,
RNAseq)

Plants respond to excess light via myriad of mechanisms. (physiological response to excess light,
transcriptomics. broad overview to next 2 subsections)



\subsection{Physiological Responses to Excess Light}
\label{subsec:intro-resp-physio}

Upon exposure to excess light, plants immediately  begin to mitigate its detrimental effects.
(summary of photoprotective mechanisms? this could be duplication of above)

In addition to these chloroplastic photoprotective mechanisms, plants can respond to excess light on
a cellular scale. Chloroplast avoidance movement, the movement of chloroplasts parallel to the
incident angle of light, decreases the amount of absorbed light
\autocite{kasahara_chloroplast_2002}. Transcriptional induction of heat shock proteins, antioxidant
scavenging and Photooxidative damage repair occur following excess light, and help to minimise oxidative
damage \autocite{niyogi_photoprotection_1999}. Production of anthocyanin, a protective class of
pigments, is induced by the production of reactive oxygen species due to excess light
\autocite{vanderauwera_genome-wide_2005}. Together, these responses serve to reduce the Photooxidative damage
of cells and tissues due to excess light.



\subsection{Transcriptional Responses to Light Stress}
\label{subsec:intro-resp-txome}

Transcriptomics, or the global study of gene expression (discussed further below), has been used to
study the response of plants to excess light. 

The rapid response of the \textit{Arabidopsis} transcriptome to changes in lighting has been
demonstrated.  \textcite{rossel_global_2002} demonstrated 

Acclimation to excess light conditions affects the steady-state level of expression of several
classes of genes. In the cyanobacterium \textit{Synechocysitis} sp. PCC 6803, genes involved in light
capture are down-regulated and homologues of heat-shock proteins up-regulated after 15 hours of
excess light treatment \autocite{hihara_dna_2001}. In rice, similar transcriptional down-regulation
of light harvesting and up-regulation of photoprotection after 24 or 72 hours of excess light
treatment has been observed \autocite{murchie_acclimation_2005}. Using quantitative PCR,
\textcite{gordon_systemic_2013} have show that repeated high light treatments lead to acclimation and
reduced induction of high light responsive transcripts.

Expression of \underline{A}SCORBATE \underline{P}ERO\underline{X}IDASE 2 (APX2), ELIP, HSP, LCHB, is
induced by hot excess light. These genes can be used as markers\ldots (this should state the
rationale behind qPCR marker genes, from gordon, rossell, estavillio etc).

Many genes' induction by excess light is dependent on light quality.
\textcite{gordon_systemic_2013} demonstrate that hot excess light induced expression of <+ excess
light responsive genes+> was dependent on light quality in \textit{Arabidopsis}. (expand this with
other examples) 

These transcriptomic responses to light overlap with response to
drought, pathogen or oxidative stresses as well as hormone response, reiterating the requirement of
stresses to be considered in combination. No genome-wide study of combinations of stresses has, to
my knowledge, been published. (this should rebut the studies above, criticising their
single-stress approaches in light of atkinson, mittler, jansson and karpinski)



\subsection{Response of Plants to Field Conditions}
\label{subsec:intro-resp-field}

Under field conditions, plants respond to the combination of stresses they naturally encounter, as
opposed to pre-determined combinations hypothesised to provoke a mechanism or response. (state why
this is cool, with reference to latest karpinkski paper)

Plants grown in the field show markedly different phenotypes and responses to those grown in the
lab. This includes metabolic profiles \autocite{jankanpaa_metabolic_2012}, transcript abundances
\autocite{mishra_arabidopsis_2012, karpinksi papers on apx2 and latest one}, photosynthetic
differences \autocite{mishra_arabidopsis_2012}, and survival and fecundity
\autocite{kulheim_rapid_2002}. (flesh out each of these, might make this 2 paragraphs:
transcriptomics, and everything else).

Several authors report overlap between genes constitutive expressed in field conditions, and those
expressed upon treatment with excess light. (ELIPs, Jansen and Karpinski.
\autocite{mishra_arabidopsis_2012, }


%%%%%%%% Misc %%%%%%%%%

\section{Misc/todo/not sure where to stick but important points}

\todo{rebuttal to state of art in biol.\ literature}

Light quality is particularly important in studies of Photooxidative damage, as the extent of photosystem II
damage is not consistent across the visible spectrum \autocite{takahashi_solar_2010}. Photooxidative damage is
relatively more severe under light of wavelengths between 580-620nm than in the remainder of the
visible spectrum, which overlaps with a density peak in the spectral power density of fluorescent
lamps (\autoref{fig:intro-dfl-kulheimsolaroutput}).

\todo{genetic variation}
%Photoprotecion are, they in general have not examined the extent and effect of natural genetic
%variation, or combinatorial application of abiotic stresses. 


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Transcriptomics %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transcriptomics: Assessing Global Expression}
\label{sec:intro-txomics}

By studying how, when and to what extent each gene in
the genome is expressed, we can gain insight into the response to any
perturbation to a plant's environment. To study global expression, two approaches have commonly been
used: microarrays and RNAseq. Microarrays quantify expression as fluorescence from
fluorophore-tagged cDNAs complementary to specific small sections of the transcriptome. More modern
experiments use RNAseq, the high-throughput sequencing of cDNA libraries, and rely on the number of
sequence reads being proportional to the concentration of a particular transcript within the
library. RNAseq allows not only quantification, but provides the actual sequence of each transcript,
important for analyses of sequence variation such as allele specific expression.



\subsection{Theoretical Aspects of RNAseq}
\label{subsec:intro-txomics-theory}

RNA sequencing (RNAseq) is a modern method of whole-transcriptome quantification. By creating cDNA
from RNA, and sequencing libraries of this cDNA, quantification is possible. (explain each step
briefly this).

As RNAseq delivers raw sequence data, but aims to provide quantification of the expression of each
gene in the genome, further analysis is necessary. ( 

Quantitative data obtained via RNAseq must undergo statistically rigorous normalisation and
hypothesis testing. The statistical basis underlying RNAseq is an area of active research.
(\autocite{robinson})

Post-hoc analyses of expression data can be performed to glean summarised biological meaning from
genome-wide differential expression patterns. (GO analysis. Try finding papers to back up my
criticism of GO as being tautologous bollocks.)


\subsection{Advances in RNAseq Methodology}
\label{subsec:intro-txomics-advances}

Recent years have seen an explosion of improvements in published protocols and tools for the
creation and analysis of RNAseq datasets.

Higher throughput and less expensive RNAseq sequencing library preparation methods are required for
large-scale RNAseq experiments. (paraphrase some good points from intro of Kumar et al)

RNAseq analysis software with improved performance and accuracy is continually published. (subread,
latest edgeR paper, QLspline paper, etc.)



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Aims %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Thesis Aims}
\label{sec:intro-aims}


The overall goal of this project is to examine the effect of light intensity on
\textit{Arabidopsis}, within the framework of realistic combinations of abiotic stresses. Due to
technical obstacles and delays, the specific aims of this project have evolved over the year.
Originally, this project aimed to:
\begin{enumerate}
  \item Determine the response of Arabidopsis gene expression to novel dynamic light conditions.
  \item Examine the extent of genetic variation in gene expression under these dynamic light
    conditions, and elucidate gene regulation networks controlling gene expression.
  \item Explore the effect of genotype-environment interactions on gene expression under dynamic
    light conditions.
\end{enumerate}

As it became apparent that technical difficulties would prevent high-throughput RNAseq experiments
in the time available, the aims were revised to remove reliance on such techniques. Thus, the
remainder of this thesis is devoted to the examination of the following aims:
\begin{enumerate}
  \item Design and implement ``Dynamic Growth Conditions'' which mimic regional climates
  \item Select optimal software for the high-throughput study of transcriptome dynamics using
    High-throughput Sequencing, and implement a framework for generation of analysis pipelines to do
    so
  \item Determine the transcriptional response of \textit{Arabidopsis thaliana} to combinatorial
    application of abiotic stresses, using dynamic growth conditions
\end{enumerate}


